# Code for Eyelid Figure 1

## Load libraries and data
```{r}
library(Seurat)
library(dplyr)
library(ggplot2)
library(patchwork)
library(DoubletFinder)
```

## Intial quality control

### Load three datasets
```{r}
mg1.data <- Read10X(data.dir = "Data/cellranger_output/run_count_MG1_S10/outs/filtered_feature_bc_matrix")
mg2.data <- Read10X(data.dir = "Data/cellranger_output/run_count_MG2_S11/outs/filtered_feature_bc_matrix")
mg.wt.data <- Read10X(data.dir = "/Users/juliewiedemann/Google Drive/Julie/Cornea_scRNA-seq/Data/cellranger_output/WT_MG/filtered_feature_bc_matrix")
```
### Create Seurat objects
```{r}
# Initialize the Seurat object with the raw (non-normalized data).
mg1 <- CreateSeuratObject(counts = mg1.data, project = "mg1", min.cells = 6)
#16983 features across 4755 samples within 1 assay
mg2 <- CreateSeuratObject(counts = mg2.data, project = "mg2", min.cells = 6)
#17057 features across 3428 samples within 1 assay
mg.wt <- CreateSeuratObject(counts = mg.wt.data, project = "mg_wt", min.cells = 6)

```

### Assign percent mito
```{r}
mg1[["percent.mt"]] <- PercentageFeatureSet(mg1, pattern = "^mt-")
mg2[["percent.mt"]] <- PercentageFeatureSet(mg2, pattern = "^mt-")
mg.wt[["percent.mt"]] <- PercentageFeatureSet(mg.wt, pattern = "^mt-")
```

### QC violin plot
```{r}
pdf("Plots/MG1_QC_plot.pdf")
VlnPlot(mg1, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), pt.size = 0, ncol = 3)
dev.off()

pdf("Plots/MG2_QC_plot.pdf")
VlnPlot(mg2, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), pt.size = 0, ncol = 3)
dev.off()

pdf("Plots/MG_WT_QC_plot.pdf")
VlnPlot(mg.wt, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), pt.size = 0, ncol = 3)
dev.off()
```

### SuppFig1- QC violin plot after cutoffs were applied
```{r}
mg1 <- subset(mg1, subset = nFeature_RNA > 200 & nFeature_RNA < 6000 & percent.mt < 5)
#An object of class Seurat
#16983 features across 4267 samples within 1 assay
#Active assay: RNA (16983 features, 0 variable features)
mg2 <- subset(mg2, subset = nFeature_RNA > 200 & nFeature_RNA < 6000 & percent.mt < 5)
#An object of class Seurat
#17057 features across 2992 samples within 1 assay
#Active assay: RNA (17057 features, 0 variable features)
mg.wt <- subset(mg.wt, subset = nFeature_RNA > 200 & nFeature_RNA < 6000 & percent.mt < 5)

pdf("Plots/MG1_after_QC_plot.pdf")
VlnPlot(mg1, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), pt.size = 0, ncol = 3)
dev.off()

pdf("Plots/MG2_after_QC_plot.pdf")
VlnPlot(mg2, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), pt.size = 0, ncol = 3)
dev.off()

pdf("Plots/MG_WT_after_QC_plot.pdf")
VlnPlot(mg.wt, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), pt.size = 0, ncol = 3)
dev.off()
```

## Integrate three eyelid datasets

### Assign dataset labels
```{r}
mg1$id <- "MG1"
mg2$id <- "MG2"
mg.wt$id <- "MG3"
```

### Integrate three datasets into one object
```{r}
features <- SelectIntegrationFeatures(object.list = list(mg1, mg2, mg.wt))
anchors <- FindIntegrationAnchors(object.list = list(mg1, mg2, mg.wt), anchor.features = features)
mg.combined <- IntegrateData(anchorset = anchors)
```

### Run standard Seurat pipeline on integrated object
```{r}
DefaultAssay(mg.combined) <- "integrated"
# Run the standard workflow for visualization and clustering
mg.combined <- ScaleData(mg.combined, verbose = FALSE)
mg.combined <- RunPCA(mg.combined, verbose = FALSE)
mg.combined <- RunTSNE(mg.combined, reduction = "pca", dims = 1:20)
mg.combined <- RunUMAP(mg.combined, reduction = "pca", dims = 1:20)
mg.combined <- FindNeighbors(mg.combined, reduction = "pca", dims = 1:20)
mg.combined <- FindClusters(mg.combined, resolution = c(0.1, 0.2, 0.25, 0.3, 0.35, 0.4, 0.5, 0.6, 0.7))
```

## Assign cell type names
```{r}
Idents(mg.wt.combined) <- "integrated_snn_res.0.4"

new.cluster.ids <- c("Fibroblast 0", "Basal\nconjunctiva 1", "Ductal\n epithelia 2", "HF 3", "Undiff\nmeibocyte 4", " Suprabasal\nconjunctiva 5", "Prolif? HF 6", "Prolif HF 7", "Fibroblast 8", "? 9", "HF? 10", "T cells\n11", "Immune 12", "Conjunctiva\nprogenitor 13", "HF 14", "Meibocyte/Ductal\nprogenitor 15", "Muscle 16", "Diff\nmeibocyte 17", "Endothelial cells 18", "Fibroblast 19", "Ductal\n epithelia 20", "Melanocyte 21")
names(new.cluster.ids) <- levels(mg.wt.combined)
mg.wt.combined <- RenameIdents(mg.wt.combined, new.cluster.ids)
mg.wt.combined$wt_integrated_annotation <-Idents(mg.wt.combined)

Idents(mg.combined) <-'wt_integrated_annotation'
new.cluster.ids <- c('Fibroblast I', 'CJ progenitor', 'Ductal progenitor', 'HF I', 'Meibocyte progenitor', 'Suprabasal CJ epithelia', 'IRS', 'HF III (prolif)', 'Fibroblast II', 'DP', 'Cortex/cuticle', 'T cells/lymphoid', 'Macrophages/DC', 'CJ progenitor (prolif)', 'HF II', 'Meibocyte/Ductal progenitor(prolif)', 'Pericytes', 'Diff meibocyte', 'Endothelial cells', 'DS', 'Diff ductal epithelia', 'Melanocytes')

names(new.cluster.ids) <- levels(mg.combined)
mg.combined <- RenameIdents(mg.combined, new.cluster.ids)
mg.combined$final_annotation <- Idents(mg.combined)

my_levels <-  c('Fibroblast I', 'Fibroblast II',  'DS', 'DP', 'HF I', 'HF II', 'HF III (prolif)', 'IRS', 'Cortex/cuticle', 'CJ progenitor (prolif)', 'CJ progenitor', 'Suprabasal CJ epithelia', 'Meibocyte/Ductal progenitor(prolif)', 'Meibocyte progenitor', 'Diff meibocyte', 'Ductal progenitor', 'Diff ductal epithelia', 'T cells/lymphoid', 'Macrophages/DC', 'Pericytes', 'Endothelial cells', 'Melanocytes')
mg.combined.s$final_annotation <- factor(x = mg.combined.s$final_annotation, levels = my_levels)
Idents(mg.combined.s) <- "final_annotation"
```

## Perform doublet removal on each dataset

### Split object by sample

```{r}
DefaultAssay(mg.combined) <- 'RNA'
mg_sample <- SplitObject(mg.combined, split.by = 'id')
```

### MG 1
```{r}
mg_sample[["MG1"]] <- mg_sample[["MG1"]] %>% NormalizeData() %>% FindVariableFeatures() %>% ScaleData() %>% RunPCA()

sweep.res.list_mg1 <- paramSweep_v3(mg_sample$MG1, PCs = 1:20, sct = FALSE)
sweep.stats_mg1 <- summarizeSweep(sweep.res.list_mg1, GT = FALSE)
bcmvn_mg1 <- find.pK(sweep.stats_mg1)

## Homotypic Doublet Proportion Estimate -------------------------------------------------------------------------------------
homotypic.prop_mg1 <- modelHomotypic(mg_sample$MG1$final_annotation)           ## ex: annotations <- seu_kidney@meta.data$ClusteringResults
nExp_poi_mg1 <- round(0.05*nrow(mg_sample$MG1@meta.data))  ## Assuming 7.5% doublet formation rate - tailor for your dataset
nExp_poi.adj_mg1 <- round(nExp_poi_mg1*(1-homotypic.prop_mg1))

## Run DoubletFinder with varying classification stringencies ----------------------------------------------------------------
mg_sample[["MG1"]] <- doubletFinder_v3(mg_sample[["MG1"]], PCs = 1:20, pN = 0.25, pK = 0.005, nExp = nExp_poi_mg1, reuse.pANN = FALSE, sct = FALSE)
mg_sample[["MG1"]] <- doubletFinder_v3(mg_sample[["MG1"]], PCs = 1:20, pN = 0.25, pK = 0.005, nExp = nExp_poi.adj_mg1, reuse.pANN = "pANN_0.25_0.005_213", sct = FALSE)
```
### MG2
```{r}
mg_sample[["MG2"]] <- mg_sample[["MG2"]] %>% NormalizeData() %>% FindVariableFeatures() %>% ScaleData() %>% RunPCA()

sweep.res.list_mg2 <- paramSweep_v3(mg_sample$MG2, PCs = 1:20, sct = FALSE)
sweep.stats_mg2 <- summarizeSweep(sweep.res.list_mg2, GT = FALSE)
bcmvn_mg2 <- find.pK(sweep.stats_mg2)

## Homotypic Doublet Proportion Estimate -------------------------------------------------------------------------------------
homotypic.prop_mg2 <- modelHomotypic(mg_sample$MG2$final_annotation)           ## ex: annotations <- seu_kidney@meta.data$ClusteringResults
nExp_poi_mg2 <- round(0.05*nrow(mg_sample$MG2@meta.data))  ## Assuming 7.5% doublet formation rate - tailor for your dataset
nExp_poi.adj_mg2 <- round(nExp_poi_mg2*(1-homotypic.prop_mg2))

## Run DoubletFinder with varying classification stringencies ----------------------------------------------------------------
mg_sample[["MG2"]] <- doubletFinder_v3(mg_sample[["MG2"]], PCs = 1:20, pN = 0.25, pK = 0.01, nExp = nExp_poi_mg2, reuse.pANN = FALSE, sct = FALSE)
mg_sample[["MG2"]] <- doubletFinder_v3(mg_sample[["MG2"]], PCs = 1:20, pN = 0.25, pK = 0.01, nExp = nExp_poi.adj_mg2, reuse.pANN = "pANN_0.25_0.08_150", sct = FALSE)
```
### MG3
```{r}
mg_sample[["MG3"]] <- mg_sample[["MG3"]] %>% NormalizeData() %>% FindVariableFeatures() %>% ScaleData() %>% RunPCA()

sweep.res.list_mg3 <- paramSweep_v3(mg_sample$MG3, PCs = 1:20, sct = FALSE)
sweep.stats_mg3 <- summarizeSweep(sweep.res.list_mg3, GT = FALSE)
bcmvn_mg3 <- find.pK(sweep.stats_mg3)

## Homotypic Doublet Proportion Estimate -------------------------------------------------------------------------------------
homotypic.prop_mg3 <- modelHomotypic(mg_sample$MG3$final_annotation)           ## ex: annotations <- seu_kidney@meta.data$ClusteringResults
nExp_poi_mg3 <- round(0.05*nrow(mg_sample$MG3@meta.data))  ## Assuming 7.5% doublet formation rate - tailor for your dataset
nExp_poi.adj_mg3 <- round(nExp_poi_mg3*(1-homotypic.prop_mg3))

## Run DoubletFinder with varying classification stringencies ----------------------------------------------------------------
mg_sample[["MG3"]] <- doubletFinder_v3(mg_sample[["MG3"]], PCs = 1:20, pN = 0.25, pK = 0.005, nExp = nExp_poi_mg3, reuse.pANN = FALSE, sct = FALSE)
mg_sample[["MG3"]] <- doubletFinder_v3(mg_sample[["MG3"]], PCs = 1:20, pN = 0.25, pK = 0.005, nExp = nExp_poi.adj_mg3, reuse.pANN = "pANN_0.25_0.005_524", sct = FALSE)
```


## Remove doublets from dataset and finalize annotation
```{r}
singlet <- c(colnames(subset(mg_sample$MG1, idents = "Singlet")), colnames(subset(mg_sample$MG2, idents = "Singlet")),colnames(subset(mg_sample$MG3, idents = "Singlet")))

mg.combined.s <- subset(mg.combined, cells = singlet)
```

## Fig1A- UMAP with updated names
Color palette for cell types
```{r}
pal <- c("#FF474E", "#F67513", "#F9AF39", "#F8EB77",
"#DFE58A", "#BAE576", "#A1DB43", "#73C658", "#55A739",
"#C0EBEC", "#61CDD1", "#32AAAE",
"#A1DAF7", "#54BEF2", "#1090D1", "#BCA3F5", "#936CEF",
"#E19BE9", "#DE3FBC",
"#FFADCD", "#FF85B4", "#FF478E"
)
```

Plot UMAP with cell type annotation
```{r}
pdf("Plots/MG_integrated/singlet/UMAP_MG_all_clusters_annotated.pdf", height = 8, width = 8)
DimPlot(mg.combined.s, cols = pal, label = TRUE)+NoLegend()
dev.off()
```

## Fig1B- Cell type marker violin plot
Color palette for violin plot
```{r}
pal2 <- c("#FF474E", "#F9AF39", "#F8EB77",
"#DFE58A", "#A1DB43", "#73C658", "#55A739",
"#C0EBEC", "#32AAAE",
"#54BEF2", "#936CEF",
"#E19BE9", "#DE3FBC",
"#FFADCD", "#FF85B4", "#FF478E"
)
```

Cell type marker violin plot
```{r}
pdf("Plots/MG_integrated/singlet/StackedVlnPlot_MG_integrated_markers_singlet.pdf", height =12, width =9)
VlnPlot(mg.combined.s, features = c("Col1a1", "Acta2", "Corin", "Lhx2",  "Mki67", "Krt73", "Krt35", "Cldn4", "Krt13", "Ly6d", "Plin2", "Cd3d", "C1qa", "Rgs5", "Cdh5",  "Mlana"), same.y.lims =TRUE, stack=TRUE, flip =TRUE, cols = pal2)+NoLegend()+xlab("")
dev.off()
```

## Fig1C- Cell cycle scoring

```{r}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

mg.combined.s <- CellCycleScoring(mg.combined.s, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

pdf("/Plots/MG_integrated/singlet/UMAP_MG_integrated_CellCycleScoring_singlet.pdf")
DimPlot(mg.combined)+NoLegend()
dev.off()
mg.combined.s$cellcycle <- Idents(mg.combined.s)
```

### Dataframe with cell cycle scoring proportion by cell type

```{r}
cc <- melt(table(mg.combined.s$final_annotation, mg.combined.s$cellcycle))

t <- as.numeric(table(mg.combined.s$final_annotation))

cc$proportion <- ifelse(cc$Var.1 == levels[18], cc$value/t[18], ifelse(cc$Var.1 == levels[19], cc$value/t[19],
  ifelse(cc$Var.1 == levels[20], cc$value/t[20],  ifelse(cc$Var.1 == levels[21], cc$value/t[21],
  ifelse(cc$Var.1 == levels[1], cc$value/t[1], ifelse(cc$Var.1 == levels[2], cc$value/t[2],ifelse(cc$Var.1 == levels[3], cc$value/t[3],
  ifelse(cc$Var.1 == levels[4], cc$value/t[4], ifelse(cc$Var.1 == levels[5], cc$value/t[5], ifelse(cc$Var.1 == levels[6], cc$value/t[6],
  ifelse(cc$Var.1 == levels[7], cc$value/t[7], ifelse(cc$Var.1 == levels[8], cc$value/t[8], ifelse(cc$Var.1 == levels[9], cc$value/t[9],
  ifelse(cc$Var.1 == levels[10], cc$value/t[10], ifelse(cc$Var.1 == levels[11], cc$value/t[11], ifelse(cc$Var.1 == levels[12], cc$value/t[12],
  ifelse(cc$Var.1 == levels[13], cc$value/t[13], ifelse(cc$Var.1 == levels[14], cc$value/t[14], ifelse(cc$Var.1 == levels[15], cc$value/t[15], ifelse(cc$Var.1 == levels[16], cc$value/t[16], ifelse(cc$Var.1 == levels[17], cc$value/t[17], cc$value/t[22])))))))))))))))))
```

### Bar chart to show changes in cell cycle proportions by cell type

```{r}
pdf("Plots/MG_integrated/Barplot_proportion_MG_integrated_CellCycleScoring.pdf", height = 4, width = 15)
cc$Var.1 <- as.factor(cc$Var.1)
cc$Var.1 <- factor(x = cc$Var.1, levels = levels)

ggplot(cc, aes(x= Var.1, y=proportion, fill =Var.2))+
geom_bar(width = .75, stat = "identity") +
  xlab('Cluster') + ylab('Proportion of cells') + labs(fill='')+ theme_bw()
dev.off()
```

```{r}
save(mg.combined.s, file = "Data/MG_Integrated_dataset.RData")
```
